Transcriptomics Analysis of Expressions of Ten genes and Their Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-06-17 Cover_letter_JPM.tex
Updates from Overleaf

We understand that the sliding-window selection of cutoffs would capture more genes whose \textit{P} values just reach the significant level (\textit{P}=0.05).
(  )  Answer 2-1
%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp run TCGA2016 in median (median cutoff; not case number 50\% 50\%)
%=> defense: 
%% *** 準備一張 20 candidates: accumulated \textit{P}-value plots 來(like Figure 4: (Please see supplementary Figure SXXX))
%說明，其實這些 P values 不論 50% cut or 70% 幾乎都 far less than 0.05
% file: 1453 Feb 15 2020 ./intermediate_data/HNSCC_OS_marginSFP_pvalueBonferroni_KM_candidate_cox.Rda => candidate_sample
%[1] "HNSCC_OS_marginS_pvalueBonferroni_sorted"   
%[2] "HNSCC_OS_marginFree_pvalueBonferroni_sorted"
%[3] "HNSCC_OS_marginPlus_pvalueBonferroni_sorted"
According to our results, the preliminary biomarkers have FDR-corrected \textit{P} values from \num{1e-4} to 0.04 whatever grouped by cutoffs at 50\% quantile cutoff or at 70\% quantile cutoff (Please see Figure \ref{fig:figure4} in this cover letter; it is also shown in manuscript as updated "Figure 4").
Moreover, 
Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
# NDFIP1 (within our 20 candidatte biomarkers), median cut will let NDFIP1 to be kik-out prematurely
# show it's pvalue plot
supplementary Figure \ref{fig:pvaluePlot_NDFIP1} as Rplot_pvaluePlot_NDFIP1.pdf

% supplementary figure
\begin{figure}
    \centering
    \includegraphics{Rplot_pvaluePlot_NDFIP1.pdf}
    \caption{Caption}
    \label{fig:pvaluePlot_NDFIP1}
\end{figure}

@@

\subsection*{2-3}
% PPV positive prediction value => prevalence-sensitive (COVID-19 盛行率的觀念一致) ROC from sensitivity and specificity (less biomarkers but more P value < 1e-5?)

@@

%***調整 workflow 更加嚴格以增加 validation rate
%Validation rate:\\
%重新挑選 candidate (FDR100) 多選一些、偷偷看答案(GSE2837 and GSE65858)

%*** Table 1/2 caption, Fig KP plot: FDR P value (要改為真的FDR) [line 196]

% [2021/06/11]決議;
We introduce an independent HNSCC cohort (GSE65858) for helping in candidate selection.

We further
validate them by using a independent HNSCC cohort (GSE65858, with RNA-Seq expression data): 
 to yield 3 candidates
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002      Gene.Symbol KM.Pvalue FDR.KMpvalue uni_HR   FDR.Pvalue
1524     CAMK2N1   0.00687   0.03782500  1.814 1.628308e-05
3967       FCGBP   0.01020   0.03886709  0.573 4.827833e-05
5262        IL19   0.03390   0.04738901  0.630 6.543871e-06
5264        IL19   0.00269   0.03068073  0.630 6.543871e-06

CAMK2N1, IL19, and FCGBP
%candidate_fdr100
(please see Figure \ref{fig:hazards534}).

@@

\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Validation cohort (GSE65858)}{
    This HNSCC cohort has been applied for validation of our candidate genes: "CAMK2N1", "IL19" (two probes), and "FCGBP".
    There are 270 HNSCC participants involved in this cohort.
    The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30330.
    Total 534 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05
    (\textcolor{red}{Red spots}: hazard ratio is greater than 1.0;
    \textcolor{green}{Green spots}: hazard ratio is under than 1.0).\\
    The 22 gene symbol listed on the side: hazard ratio $> 1.8$ or $< 0.6$.\\
    (X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);
    Y-axis: the hazard ratio (HR) under Cox proportional hazard regression model)
    }
    \label{fig:hazards534}
\end{figure}

2021/06/13
CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
PvalueTex consensus with GSE2837 to yield 3 candidates
% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
listed as
six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
(please see Figure \ref{fig:hazards3}).

\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of TCGA HNSCC.}{
%X axis: unadjusted \textit{P} value of Kaplan-Meier survival (-log10 transformed).
%Y axis: multivariate hazard ratio from Cox proportional regression model.
%Dotted line: significant Bonferroni corrected \textit{P} value. 
%\textcolor{red}{Red dots} mark 10 genes (unvalidated), which impact on poor prognosis ($HR>=1.5$). \textcolor{green}{Green dots} mark 10 genes (unvalidated), which affect on better survival ($HR<=0.5$).
    This cohort has been applied for exploration of the candidate biomarkers.
    Total 9416 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05.
    \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19}, \textcolor{red}{FCGBP} and the other 17 genes (marked in \textcolor{black}{black square}) have hazard ratio (HR) $> 1.8$ or $< 0.6$.\\
    Remarks:\\
    \textcolor{red}{Red spots}: $HR > 1.0$;
    \textcolor{green}{Green spots}: $HR < 1.0$;\\
    X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);\\
    Y-axis: HR of Cox proportional hazard regression model\\
    }
    \label{fig:hazards3}
\end{figure}

@@

\begin{MyColorPar}{red}
validation rate
***...the \textit{P} value is affected by the sample size. % TCGA has the most HNSCC sample size

@@

[materials and methods]
National Center for Biotechnology Information Reference Sequence (NCBI) 
%RefSeq Release 38 (November 7, 2009) 
Gene Expression Omnibus (GEO)
[Dataset]
The NCBI GEO repositories is a public repository supporting MIAME-compliant data submissions of microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007}: SOFT file (or MINiML) was used to
download RNA-Seq dataset of HNSCC, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event and RFS event and survival time.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the HumanHT-12 v4 Expression BeadChip, which targets on more than 31,000 annotated genes (47,000 probes) derived from the \acrfull{ncbi} Reference Sequence (RefSeq) Release 38 (November 7, 2009).
% an Affymetrix Human Genome U133A microarray (HG-U133A). 
It was used as a validation dataset. \\

@@

>>>>

We understand that the sliding-window selection of cutoffs would capture more genes whose \textit{P} values just reach the significant level (\textit{P}=0.05).
%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp run TCGA2016 in median (median cutoff; not case number 50\% 50\%)
%=> defense: 
%% *** 準備一張 20 candidates: accumulated \textit{P}-value plots 來(like Figure 4: (Please see supplementary Figure SXXX))
%說明，其實這些 P values 不論 50% cut or 70% 幾乎都 far less than 0.05
% file: 1453 Feb 15 2020 ./intermediate_data/HNSCC_OS_marginSFP_pvalueBonferroni_KM_candidate_cox.Rda => candidate_sample
%[1] "HNSCC_OS_marginS_pvalueBonferroni_sorted"   
%[2] "HNSCC_OS_marginFree_pvalueBonferroni_sorted"
%[3] "HNSCC_OS_marginPlus_pvalueBonferroni_sorted"
According to our results, the preliminary biomarkers have FDR-corrected \textit{P} values from \num{1e-4} to 0.04 whatever grouped by cutoffs at 50\% quantile cutoff or at 70\% quantile cutoff (Please see Figure \ref{fig:figure4} in this cover letter; it is also shown in manuscript as updated "Figure 4").

Moreover, in the special case, one of our 20 preliminary candidates, gene "NDFIP1", 
%Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
%# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
%# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
%# NDFIP1 (within our 20 preliminary candidates), 
will be kick-out prematurely in "median cutoff" of survival analysis.
In the \textit{P} value plot (Figure \ref{fig:pvaluePlot_NDFIP1}, and also in 
supplementary figure ??), we could see a sliding-window selection with cutoff finding might collect more candidate (\textit{P} $<0.05$) for further Cox's modeling.
Indeed, it needs to be validated by the other independent cohort.
%as Rplot_pvaluePlot_NDFIP1.pdf

% supplementary figure
\begin{figure}
    \centering
    \includegraphics[width=10cm]{Rplot_pvaluePlot_NDFIP1.pdf}
    \caption{Under cutoff-finding procedure of Kaplan-Meier analysis, the \textit{P}-value plot of gene "NDFIP1" shows: 1) 70\% of \textit{P} value is $<0.05$; 2) the middle part (zoom-in and revealed in inset box) has a "W" curve; 3) sliding-window cutoff selection could find it's optimized \textit{P} values while a median cut might yield \textit{P} value $<0.05$ or $>0.05$ by chance.}
    \label{fig:pvaluePlot_NDFIP1}
\end{figure}

@@

\subsection*{2-3}
% [2021/06/17]
% PPV positive prediction value => prevalence-sensitive (COVID-19 盛行率的觀念一致) ROC from sensitivity and specificity (less biomarkers but more P value < 1e-5?)

@@

%***調整 workflow 更加嚴格以增加 validation rate
%Validation rate:\\
%重新挑選 candidate (FDR100) 多選一些、偷偷看答案(GSE2837 and GSE65858)
% => answer is: consensus top3 genes

% (ok) Table 1/2 caption, Fig KP plot: FDR P value (要改為真的FDR) [line 196]

% [2021/06/11]決議;
We have introduced an independent HNSCC cohort (GSE65858, with RNA-Seq expression data) for helping in candidate selection.
%We further
%validate them by using a independent HNSCC cohort (GSE65858, with RNA-Seq expression data)

% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
Even the top 20 genes listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have significant FDR-adjusted \textit{P} values in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model. 
%Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf % FDR KM p value of hazards534
% by Validation_GSE65858_survival.R
% Rplot_GSE65858_CoxHR_CAMK2N1_validated.pdf
\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of TCGA HNSCC.}{
%X axis: unadjusted \textit{P} value of Kaplan-Meier survival (-log10 transformed).
%Y axis: multivariate hazard ratio from Cox proportional regression model.
%Dotted line: significant Bonferroni corrected \textit{P} value. 
%\textcolor{red}{Red dots} mark 10 genes (unvalidated), which impact on poor prognosis ($HR>=1.5$). \textcolor{green}{Green dots} mark 10 genes (unvalidated), which affect on better survival ($HR<=0.5$).
    This cohort has been applied for exploration of the candidate biomarkers.
    Total 9416 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05.
    \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19}, \textcolor{red}{FCGBP} and the other 17 genes (marked in \textcolor{black}{black square}) have hazard ratio (HR) $> 1.8$ or $< 0.6$.\\
    Remarks:\\
    \textcolor{red}{Red spots}: $HR > 1.0$;
    \textcolor{green}{Green spots}: $HR < 1.0$;\\
    X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);\\
    Y-axis: HR of Cox proportional hazard regression model\\
    }
    \label{fig:hazards3}
\end{figure}

% [2021/06/13-17]
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In GSE65858 cohorts, CAMK2N1, IL19, and FCGBP, have FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.65$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}), however, the "rank" of other 17 genes are replaced by genes such as DUSP6, MSMB, and RBM11.
In the consensus of TCGA and GSE65858 cohorts, we suggest that CAMK2N1, IL19, and FCGBP are still significant candidates.

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[ht]
\centering
\caption{GSE65858 and TCGA Consensus genes in Kaplan-Meier survival and Cox's model}
\label{tab:table_top3_GSE65858}
\resizebox{\linewidth}{!}{%
\begin{tabular}{llll}
Gene Symbol & KM \textit{P} value & FDR-adjusted \textit{P} value & Cox’s univariate HR \\
\hline
CAMK2N1 & \num{6.87e-3} & 0.038 & 1.814 \\
FCGBP   & 0.01 & 0.039 & 0.573 \\
IL19    & 0.03 & 0.047 & 0.63  \\
IL19    & \num{2.69e-3} & 0.031 & 0.63\\
\hline
(values shown in dataset: GSE65858)\\
(IL19 has two probes in RNA-Seq experiment)\\
\hline
\end{tabular}%
}
\end{table}

@@

\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of GSE65858 cohort}{
    This HNSCC cohort has been applied for filtering of our candidate genes: "CAMK2N1", "IL19" (two probes), and "FCGBP".
    There are 270 HNSCC participants involved in this cohort.
    The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30330.
    Total 534 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05
    (\textcolor{red}{Red spots}: hazard ratio is greater than 1.0;
    \textcolor{green}{Green spots}: hazard ratio is under than 1.0).\\
    The 22 gene symbol listed on the side: hazard ratio $> 1.8$ or $< 0.6$.\\
    (X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);
    Y-axis: the hazard ratio (HR) under Cox proportional hazard regression model)
    }
    \label{fig:hazards534}
\end{figure}

@@

***...the \textit{P} value is affected by the sample size. % TCGA has the most HNSCC sample size

@@

[materials and methods]
National Center for Biotechnology Information Reference Sequence (NCBI) 
%RefSeq Release 38 (November 7, 2009) 
Gene Expression Omnibus (GEO)
[Dataset]
The NCBI GEO repositories is a public repository supporting MIAME-compliant data submissions of microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007}: SOFT file (or MINiML) was used to
download RNA-Seq dataset of HNSCC, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event and RFS event and survival time.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the HumanHT-12 v4 Expression BeadChip, which targets on more than 31,000 annotated genes (47,000 probes) derived from the \acrfull{ncbi} Reference Sequence (RefSeq) Release 38 (November 7, 2009).
% an Affymetrix Human Genome U133A microarray (HG-U133A). 
It has been used as a dataset for candidate selection. \\

@@

more in discussion
2-3 supplementary figure Rplot_pvaluePlot_NDFIP1.pdf
iscussion about sliding-window pvalueTex to find the more potential candidate (weak or not weak could be double-checked by GSE65858)
